---
title: "Practical Machine Learning Project"
author: "Famibelle Medhi"
date: "Sunday, November 23, 2014"
output: pdf_document
---

The goal of the project is to predict the manner in which the tech geeks did the exercise. This is the "classe" variable in the training set.

```{r}
setwd("~/Coursera/data/Practical Machine Learning")
install.packages("e1071")
install.packages("caret")

pml_write_files = function(x){
    n = length(x)
    for(i in 1:n){
        filename = paste0("problem_id_",i,".txt")
        write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
    }
}
```
# Load the data

```{r}
if (!file.exists("pml-training.csv")) {
    fileURL <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
    message(paste("Please Wait! Download...", fileURL, "..."))
    download.file(fileURL, destfile = "pml-training.csv")
}

if (!file.exists("pml-testing.csv")) {
    fileURL <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
    message(paste("Please Wait! Download...", fileURL, "..."))
    download.file(fileURL, destfile = "pml-testing.csv")
}
```

# Clean the data

```{r}
pml_testing <-  read.csv("pml-testing.csv" , na.strings = c("", "NA"))
pml_training <- read.csv("pml-training.csv", na.strings = c("", "NA"))
```
# Remove all the columns filled with NAs.

```{r}
ColsWithNAs <- colSums(is.na(pml_training)) == 0
pml_training <- pml_training[, ColsWithNAs]
pml_testing  <- pml_testing[ , ColsWithNAs]
```

The  `X`, `user_name`, `raw_timestamp_part_1`, `raw_timestamp_part_2`, `cvtd_timestamp`, `new_window`, `num_window` are not part of the predictors and then can be removed from the model

```{r}
notPredictors <- 1:7
pml_training <- pml_training[, -notPredictors]
pml_testing  <- pml_testing[ , -notPredictors]
```

#BootStrap
use 75% of the data to train the model and 25% to test the model. That's give us a subset of the data that is just for training `training` and a subset of the data that is just for testing `testing`
```{r}
library(caret)
inTrain = createDataPartition(pml_training$classe, p = 0.75, list = FALSE)
training = pml_training[inTrain,]   # subset of the data for training
testing  = pml_training[-inTrain,]  # subset of the data for testing
```

# Fit a model
Now we can fit a model. We are going to use the `train` from that caret package. We are going to predict `classe` from the dataset `training` created above. We are going to use the GBM method.
```{r}
set.seed(1974)
#load("ModelFit.Rdata")
modelFit <- train(classe ~ . , data=training, method="gbm")
```

```{r}
modelFit
```

```{r}
modelFit$finalModel
```

We are going to predict the testing data `testing` by using the predict command based on the `modelFit` calculated above
```{r}
predictions <- predict(modelFit, newdata=testing)
```

# calculating the confusion matrix
We are going to run the confusion matrix from the data calculated from our model fit `predictions` against the actual results from the testing data `testing`
```{r}
confusionMatrix.PML <- confusionMatrix(predictions,testing$classe)
confusionMatrix.PML
```

**Accuracy is `r confusionMatrix.PML$overall[1]`**

# Now the predictions
Remember that the goal of the project was to predict the manner in which the tech geeks did the exercise. So we apply the model to test data `pml_testing`

```{r}
answers <- predict(modelFit,newdata=pml_testing)
pml_write_files(answers)
answers
```